---
title: "Financial Time Series"
output: html_notebook
params:
  tickers: !r c("AAPL", "MSFT", "AMD", "INTL", "GE", "GM", "SHLD")
  from_date: !r as.Date("2018-01-01")
  ar: 2
  p: 2
  q: 2
  s: !r c(5, 20)
  corr_prior: 4
  periods_to_predict: 30 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(magrittr)
library(ggplot2)
library(rethinking)
library(bayesplot)
library(tidybayes)
library(rstan)
library(ggplot2)
library(forcats)
library(purrr)
library(tidyr)
library(abind)
library(lubridate)
library(pdfetch)
```

```{r define_tickers}
tickers <- c(params$tickers, "SPY", "^VIX")
from <- params$from_date
leave_out <- params$periods_to_predict
```

```{r vis_prices}

data <- pdfetch_YAHOO(tickers, fields="adjclose", from=from, to=Sys.Date()) %>% 
  data.frame() %>%
  mutate(period = 1:n()) %>%
  gather(key="series", value="y", -period)

data %>%
  ggplot(aes(x=period, y=y, color=series)) +
  geom_line() +
  scale_y_log10() +
  theme_minimal()
```


```{r prepare_data}
prices_gathered <- data %>% 
  arrange(period, series) %>% 
  mutate(id = 1:n(), weight=1)

predictors <- matrix(0, ncol=1, nrow=max(data$period) - leave_out - 1)

stan_data <- compose_data(
                          # The data
                          prices_gathered %>% 
                            dplyr::select(-id) %>% 
                            dplyr::filter(period < max(data$period) - leave_out), 
                          N_periods = max(data$period) - leave_out - 1,
                          x = predictors,
                          N_features = ncol(predictors),
                          # Parameters controlling the model 
                          periods_to_predict = leave_out, 
                          corr_prior = params$corr_prior,
                          p = params$p, 
                          q = params$q, 
                          ar = params$ar, 
                          period_scale = (max(data$period) / (as.numeric(Sys.Date() - from) / 365)) * 8, 
                          s = array(params$s, dim=length(params$s)),
                          N_seasonality = length(params$s),
                          cyclicality_prior = 250 * 5, 
                          .n_name = n_prefix("N"))
```


```{r model}
model <- stan_model("./mbsts.stan")
```

```{r sample}
samples <- sampling(model, data=stan_data, chains=2, iter=2000, cores=4,
                    #control=list(adapt_delta=0.9, max_treedepth=15), 
                    init="0"
                    ) %>% recover_types(prices_gathered)
```

# Inspect Model

## Prediction Accuracy

### Residuals 

```{r recover_y_hat}
y_hat <- samples %>% gather_draws(log_y_hat[id]) %>%
  mutate(.value = exp(.value) - 1)
```

```{r residuals}
y_hat %>% 
  dplyr::select(id, .value) %>%
  inner_join(prices_gathered) %>% 
  mutate(
    error = .value - y
  ) %>%
  group_by(id, series) %>%
  summarize(
    mean_error = mean(error),
    rmse = sqrt(mean(error^2)), 
    sd_y_hat = sd(.value)
  ) %>% 
  dplyr::select(series, mean_error, rmse, sd_y_hat) %>%
  gather(key="metric", value="value", -series, -id) %>% 
  ggplot(aes(x = value, fill=series)) +
  geom_histogram(bins=40, position="identity", alpha = .3) + 
  facet_grid(series ~ metric, scales="free") +
  theme_minimal() +
  ggtitle("Residuals")
```

```{r visualize_predictions}
prices_gathered %>% dplyr::select(id, y, series) %>%
  inner_join(y_hat %>% 
               group_by(id) %>%
               summarize(y_hat = mean(.value))) %>% group_by(series) %>% 
  ggplot(aes(x = y, y=y_hat)) +
  geom_point(size=0.2, alpha=0.2) +
  facet_wrap(~ series) +
  scale_x_continuous(limits=c(0, NA)) + 
  scale_y_continuous(limits=c(0, NA)) + 
  annotate("segment", x = 0, y = 0, xend=max(data$y), yend=max(data$y), color="red", size=0.5, alpha=0.5) +
  theme_minimal() + 
  ggtitle("Prices vs. Mean Predictions")
```

### Recovered Time Series

```{r recover_time_series}
periodic_price_hat <- samples %>% gather_draws(log_prices[period, series])
```

```{r time_series_by_chain}
mean_periodic_price_hat <- periodic_price_hat %>%
  dplyr::filter(is.finite(.value)) %>%
  group_by(period, series) %>%
  summarize(
    y_hat = exp(mean(.value))
  ) %>%
  mutate(.chain = as.character("mean_hat"))

periodic_price_hat %>%
  dplyr::filter(is.finite(.value)) %>% 
  group_by(period, series, .chain) %>%
  summarize(
    y_hat = exp(mean(.value))
  ) %>%
  mutate(.chain = as.character(.chain)) %>%
  bind_rows(prices_gathered %>% mutate(.chain="gen") %>% rename(y_hat=y), mean_periodic_price_hat) %>% 
  mutate(alpha = case_when(
    .chain %in% c("gen", "mean_hat") ~ .8, 
    TRUE ~ .2
  )) %>%
  ggplot(aes(x = period, y = y_hat, color=series, linetype=.chain, alpha=I(alpha))) +
  geom_line(size=0.5) +
  theme_minimal() +
  ggtitle("Reconstruction of Time Series By Chain")
```

### Forecast

```{r gather_price_forecast}
price_forecast <- samples %>% 
  gather_draws(log_prices_hat[period, series]) 
```

```{r visualize_price_forceast}
true_future <- data %>%
  dplyr::filter(period > max(data$period) - leave_out) %>%
  mutate(period = period - (max(data$period) - leave_out))
  

price_forecast %>%
  dplyr::filter(is.finite(expm1(.value))) %>%
  ungroup() %>%
  group_by(series, period) %>%
  summarize(y = expm1(mean(.value))) %>%
  mutate(chain = "mean_hat") %>% 
  bind_rows(true_future %>% mutate(chain="gen")) %>%
  ggplot(aes(x=period, y=y, color=series, linetype=chain)) +
  geom_line() + 
  geom_lineribbon(alpha=0.3, size=0.5, data=price_forecast %>% 
                    dplyr::filter(is.finite(.value)) %>%
                    mutate(y=expm1(.value), chain="median") %>%
                    ungroup() %>%
                    dplyr::select(series, period, y, chain) %>% 
                    group_by(period, series, chain) %>% 
                    median_qi(y, .width=c(0.2, 0.3, 0.66))
                    ) +
  scale_fill_brewer("Confidence", palette="Blues") +
  theme_minimal() +
  scale_y_continuous(limits=c(0, 500), oob=function(x, limits) case_when(
    x > limits[2] ~ limits[2], 
    x < limits[1] ~ limits[1], 
    TRUE ~ x
  )) + 
  ggtitle("Forecasted Prices")
```

### Combined Plot

```{r combined_plot}
bind_rows(prices_gathered %>% 
            mutate(.chain="Actual") %>% 
            rename(y_hat=y), 
          mean_periodic_price_hat,
          price_forecast %>%
            dplyr::filter(is.finite(expm1(.value))) %>%
            ungroup() %>%
            mutate(period = period + max(mean_periodic_price_hat$period)) %>%
            group_by(series, period) %>%
            summarize(y_hat = expm1(mean(.value))) %>%
            mutate(.chain = "forecast")) %>% 
  #dplyr::filter(.chain %in% c("forecast")) %>% 
  ggplot(aes(x = period, y = y_hat, color=series, linetype=.chain)) +
  geom_lineribbon(alpha=0.2, size=0.5, data=price_forecast %>% 
                  dplyr::filter(is.finite(.value)) %>%
                  ungroup() %>%
                  mutate(
                    y_hat=expm1(.value), 
                    .chain="median",
                    period = period + max(mean_periodic_price_hat$period)
                  ) %>%
                  dplyr::select(series, period, y_hat, .chain) %>% 
                  group_by(period, series, .chain) %>% 
                  median_qi(y_hat, .width=c(0.1, 0.2, 0.3))
                  ) +
  geom_line(size=0.5, alpha=0.8) +
  scale_fill_brewer("Confidence", palette="Blues") +
  scale_y_continuous(limits=c(0, 300), oob=function(x, limits) case_when(
    x > limits[2] ~ limits[2], 
    x < limits[1] ~ limits[1], 
    TRUE ~ x
  )) +
  theme_minimal() +
  ggtitle("Reconstructed Time Series and Forecast vs. Actual")
```

```{r other_combined_plot}
bind_rows(data %>% mutate(.chain="Actual") %>% rename(y_hat=y), mean_periodic_price_hat) %>% 
  dplyr::filter(.chain %in% c("Actual", "mean_hat")) %>% 
  ggplot(aes(x = period, y = y_hat, color=series, linetype=.chain)) +
  geom_line(size=0.5) +
  geom_line(data=price_forecast %>% 
    ungroup() %>% 
    dplyr::filter(is.finite(.value)) %>% 
    group_by(series, period) %>%
    summarize(y_hat = exp(mean(.value))) %>% 
    mutate(
      period = period + (max(data$period) - leave_out),
      .chain = "forecast"
    ) %>%
    group_by(series), 
    alpha=0.6, color="darkblue", size=0.1) +
  scale_fill_brewer("Confidence", palette="Blues") +
  scale_y_log10(limits=c(1, 300)) +
  #scale_y_continuous(limits=c(0, 300)) +
  theme_minimal() +
  ggtitle("Reconstructed Time Series and Forecast vs. Actual", 
          subtitle="Predicting 120 days of closing prices from 60 days of data")
```

## Components 

### Visualize All Components

```{r recover_components}
combined_components <- samples %>%  gather_draws(delta[period, series], tau_s[seasonality][period, series], omega[period, series], epsilon[period, series], xi[period, series])  %>% 
  ungroup() %>% 
  mutate(
    component = case_when(
      .variable == "delta" ~ "trends", 
      .variable == "epsilon" ~ "innovations", 
      .variable == "tau_s" ~ paste("seasonality", seasonality, sep="_"), 
      .variable == "omega" ~ "cyclicality", 
      .variable == "xi" ~ "predictors"
    )
  )

components_mean <- combined_components %>%
  ungroup() %>%
  group_by(period, series, component) %>%
  summarize(.value = mean(.value)) %>% 
  mutate(.chain = "mean_hat")

components_mean %>%
  ggplot(aes(x = period, y = .value, color=component)) +
  facet_grid(series ~ ., scales="free_y") +
  geom_line(size=0.2) +
  theme_minimal() +
  ggtitle("Components of Price Movements (Means)")
```

```{r view_components_by_chain}
components_by_chain <- combined_components %>%
  ungroup() %>%
  group_by(period, series, component, .chain) %>%
  summarize(.value = mean(.value)) %>%
  mutate(.chain = as.character(.chain))

bind_rows(components_by_chain, components_mean) %>%
  mutate(alpha = case_when(
    .chain == "mean_hat" ~ 1, 
    TRUE ~ 0.2
  )) %>%
  ggplot(aes(x = period, y = .value, color=component, linetype=factor(.chain), alpha=I(alpha))) +
  facet_grid(series ~ ., scales="free_y") +
  geom_line(size=0.2) +
  theme_minimal() +
  ggtitle("Components of Price Movements (By Chain)")
```

### Components in Prediction Period

```{r prediction_components}
prediction_components <- samples %>%  gather_draws(delta_hat[period, series], tau_hat_all[period, series], omega_hat[period, series], epsilon_hat[period, series])  %>% 
  ungroup() %>% 
  mutate(
    component = case_when(
      .variable == "delta_hat" ~ "trends", 
      .variable == "epsilon_hat" ~ "innovations", 
      .variable == "tau_hat_all" ~ "seasonality", 
      .variable == "omega_hat" ~ "cyclicality", 
      .variable == "xi_hat" ~ "predictors"
    )
  )

predictions_mean <- prediction_components %>%
  ungroup() %>%
  group_by(period, series, component) %>%
  summarize(.value = mean(.value)) %>% 
  mutate(.chain = "mean_hat")

predictions_mean %>%
  ggplot(aes(x = period, y = .value, color=component)) +
  facet_grid(series ~ ., scales="free_y") +
  geom_line(size=0.2) +
  theme_minimal() +
  ggtitle("Components of Price Movements (Means) in Forecast")
```

### Price Movements in Forecast

```{r price_movements_in_forecast}
prediction_components %>%
  dplyr::filter(is.finite(.value)) %>% 
  ungroup() %>%
  group_by(series, period, .draw) %>% 
  summarize(
    movement = sum(.value)
  ) %>%
  group_by(series, period) %>%
  summarize(
    mean_movement = mean(movement)
  ) %>%
  ggplot(aes(x=period, y=mean_movement)) +
  geom_line() +
  facet_grid(series ~ .) +
  theme_minimal()
```

```{r cum_price_movements_in_forecast}
prediction_components %>%
  dplyr::filter(is.finite(.value)) %>% 
  ungroup() %>%
  group_by(series, period, .draw) %>% 
  summarize(
    movement = sum(.value)
  ) %>%
  group_by(series, period) %>%
  summarize(
    mean_movement = mean(movement)
  ) %>%
  arrange(period) %>%
  mutate(mean_movement = cumsum(mean_movement)) %>% 
  ggplot(aes(x=period, y=mean_movement)) +
  geom_line() +
  facet_grid(series ~ .) +
  theme_minimal()
```

### Cyclicality

```{r cyclicality}
cylicality_hat <- samples %>% gather_draws(lambda[series], rho[series])
```

```{r view_cyclicality}
cylicality_hat %>% 
  ggplot(aes(x = .value)) + 
  geom_histogram(bins=60) +
  facet_grid(series ~ .variable, scales="free") +
  scale_x_log10() +
  theme_minimal() + 
  ggtitle("Estimates of Cycle Frequency and Damping")
```

## Correlations 

### Trend Correlations

```{r recover_return_correlations}
trend_corr_hat <- samples %>% gather_draws(trend_corr[series1, series2])

true_trends <- omega_trend %>%
  data.frame() %>%
  set_colnames(1:ncol(omega_trend)) %>%
  mutate(series1 = 1:n()) %>%
  gather(key="series2", value=".value", -series1)
```

```{r show_trend_corr}
trend_corr_hat %>%
  dplyr::filter(series1 > series2) %>%
  ggplot(aes(x = .value)) +
  geom_histogram(bins=100) +
  geom_vline(data=data.frame(xintercept=0), aes(xintercept=xintercept), color="red", size=0.2) +
  facet_grid(series1 ~ series2, scales="free_y") +
  scale_x_continuous(limits=c(-1, 1)) +
  theme_minimal() + 
  ggtitle("Trend Correlations")
```

### Innovation Correlations

```{r recover_innovation_correlations}
innovation_corr_hat <- samples %>% gather_draws(innovation_corr[series1, series2])
```

```{r show_innovation_corr}
innovation_corr_hat %>%
  dplyr::filter(series1 > series2) %>%
  ggplot(aes(x = .value)) +
  geom_histogram(bins=100) +
  geom_vline(data=data.frame(xintercept=0), aes(xintercept=xintercept), color="red", size=0.2) +
  facet_grid(series1 ~ series2, scales="free_y") +
  scale_x_continuous(limits=c(-1, 1)) +
  theme_minimal() + 
  ggtitle("Innovation Correlations")
```


## Coefficients and Shrinkage Effects

### Visualize Coefficients

```{r visualize_coefficients}
coefficients_hat <- samples %>% gather_draws(beta_ar_c[feature, series], beta_xi[feature, series], 
                                             beta_p_c[feature, series], beta_q_c[feature, series])

coefficients_hat %>%
  ggplot(aes(x = .value)) +
  geom_histogram(bins=60) +
  facet_grid(series ~ .variable + feature, scales="free_x") +
  theme_minimal() +
  ggtitle("Estimated Coefficients")
```


## Volatilities and Intercepts 

### Visualize Volatilities and Intercepts

```{r visualize_volatilities}
volatilities_hat <- samples %>% gather_draws(theta_ar[series], alpha_ar[series], theta_cycle[series], theta_season[seasonality][series], omega_garch[series]) %>%
  ungroup() %>% 
  mutate(
    component = case_when(
      .variable == "theta_season" ~ paste("theta_season", seasonality, sep="_"),
      TRUE ~ .variable
    )
  )

volatilities_hat %>% 
  ggplot(aes(x = .value)) +
  geom_histogram(bins=60) +
  facet_grid(component ~ series, scales="free_x") +
  theme_minimal() +
  ggtitle("Estimated Volatilities and Baselines")
```


## Vix Reconstruction

```{r vix}
vix_hat <- samples %>% gather_draws(vix_hat[forecast_period])

bind_rows(
  vix %>% dplyr::select(y) %>% mutate(.chain="actual"), 
  vix_hat %>% dplyr::select(y = .value, .chain) %>% mutate(.chain = as.character(.chain))
) %>% 
  ggplot(aes(x=y)) + 
  geom_density() +
  facet_wrap(~ .chain) + 
  theme_minimal()
```

# DONE